Sparse Word Embedding for Multi-domain Machine Translation

### 2019-04-23 2019/emnlp/sparse.tex
more todos

\section{Related Work \label{sec:related_work}}
\fyTodo{Add standard labels to sections}
\fyTodo{Related work goes last}
Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory. Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora. The methods in this category can be listed as following:
\fyTodo{Organize refs}
\begin{itemize}
	\item score relatedness of sentences to the target domain by Language Model \cite{P10-2041,D11-1033,P13-2119}.
	\item generating pseudo parallel data \cite{P03-1010,C16-1295,D14-1023}
\end{itemize}
On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
\begin{itemize}
	\item Trainining objective \cite{Luong2015SNMT,P16-1009,D17-1155,W17-3205,D17-1156,C18-1269}
	\item Architecture \cite{gulcehre2016monolingual,C16-1170,R17-1049,W17-4712,Biao2017CARENMT,N18-2080,W18-6313,P18-2050}.
	\item The decoding algorithm \cite{gulcehre2016monolingual,I17-2004}
\end{itemize} 
Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}. The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041,D18-1041} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$. $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively. While introducing new features in the architecture, \cite{D18-1041} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side. The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation. \cite{D18-1039} has other idea to circumvent the catastrophic interference. They supposed that parameters of each domain specialized model could be generated from a universal parameter-generator. By introducing domain embedding (or Language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.

@@

\begin{table}[H]
\fyTodo{Too small ! Use abbrevs (En, Fr, ECB, EMEA, EPPS)}
  \resizebox{\columnwidth}{!}{
\begin{tabular}{ |l|l|l|l|l|l| }
\hline
Task & Corpora & Train & Dev & Test & Vocabulary \\ \hline
\multirow{5}{*}{$En \rightarrow Fr$} & European Medicines Agency & 1090568 & 1000 & 1000 & English: 30164, French: 30398\\
 & European Central Bank & 191960 & 1000 & 1000 &\\
 & European Parliament & 2005723 & 1000 & 1000 & \\ \hline
\multirow{5}{*}{$En \rightarrow Ge$} & European Medicines Agency & 1108752 & 1000 & 1000 & English: 30159, German: 30698\\
 & European Central Bank & 111174 & 1000 & 1000 &\\
 & European Parliament & 1920209 & 1000 & 1000 & \\ \hline
\hline
\end{tabular}
}
\caption{Corpora }
\label{table:Corpora}
\end{table}
\subsection{NMT model configuration}
We evaluate our method on two state-of-the-art encoder-decoder architectures: a recurrent architecture based on attentional birectional Long Short-Term Memory (LSTM) cells and the Transformer architecture of \citet{NIPS2017_7181}.
More precisely, for the Transformer we use the same settings as described in \cite{NIPS2017_7181}. For the recurrent architecture, in order to fairly compare to the work of \citet{D18-1041}, we use a standard configuration with one layer bidirectional LSTM with a encoder hidden size of 1024; one attentional unidirectional LSTM layer with same hidden size in decoder; and word embeddings of size 512. The implementation can be found in OpenNMT-tf \cite{P17-4012}.
\fyTodo{Language pairs, BPE number, optimizer}
\fyTodo{Motivate the split}
For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions $[8,8,8,488]$, corresponding respectively to EMEA, EPPS, ECB domain and generic region. If sentence comes from domain $i$, regions for domains $j,\forall j \neq i$ will be masked as 0. We then use a dense layer of size 512 to fuse the active domain region and the generic region.

@@

Each iteration of algorithm \ref{algo:1}, the batch is doubled because it contains 2 types of word embedding for each sentence (using only generic region; using generic region and active domain region). The reason to do this is to minimize simultaneously both losses \\
$\theta^*_{s}=\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]$ \\ 
$ \theta^*_{i}=\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
$ 
We admit that this is not the optimal way. However, this procdedure make the training very simple.\\
We also take in account the problem of overfitting that could occur for small corpora. Indeed, real data are not always balanced for every domain. For instance, in our experiments, the size of corpora are very different where EPPS is largest with approximately 2 millions sentences, EMEA contains about 1 million sentences and ECB is the smallest with only 200 000 sentences for English French and 100 000 for English German. If we train each corpora with same number of batches, ECB will be trained too oftenly, that easily leads to overfitting. Therefore, we perform backpropagation on corpora with frequencies according to sizes of corpora.
\subsection{Results}
\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 &  & 25.7 \\
 \hline
 EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 &  & 11.10 \\
 \hline
 EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 &  & 11.1 \\
 \hline
 ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 &  & 7.19 \\
 \hline
 Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(3) &  &  &  &  &  &  &  & \\
 \hline
 Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  & & \\
 \hline
\end{tabularx}}
\end{center}
\caption{bpe-detokenized BLEU score in language pair English - French. Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT. Sparse(2) uses generic tag for all of test sets. Sparse(3) uses tag predicted by pretrained domain classififer}
\label{tab:1}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & & & & & & & \\
 \hline
 EMEA-finetuned & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & \\
 \hline
 Sparse & & & & & & & \\
 \hline
 Domain Control & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & & & & & & & \\
 \hline 
\end{tabularx}}
\end{center}
\caption{BLEU score for Recurrent network on language pair English - German. The notation is same as figure \ref{tab:3}}
\label{tab:4}
\end{figure}
\subsection{Word embedding analysis}

\section{Conclusions}

>>>>

\section{Related Work \label{sec:related_work}}
\fyTodo{Add standard labels to sections}
\fyTodo{Related work goes last}
\fyTodo{Compare also to Peng 2017}

Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory. Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora. The methods in this category can be listed as following:
\fyTodo{Organize refs}
\begin{itemize}
	\item score relatedness of sentences to the target domain by Language Model \cite{P10-2041,D11-1033,P13-2119}.
	\item generating pseudo parallel data \cite{P03-1010,C16-1295,D14-1023}
\end{itemize}
On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
\begin{itemize}
	\item Trainining objective \cite{Luong2015SNMT,P16-1009,D17-1155,W17-3205,D17-1156,C18-1269}
	\item Architecture \cite{gulcehre2016monolingual,C16-1170,R17-1049,W17-4712,Biao2017CARENMT,N18-2080,W18-6313,P18-2050}.
	\item The decoding algorithm \cite{gulcehre2016monolingual,I17-2004}
\end{itemize} 
Beside these works, we could also consider works in other tasks such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}. The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041,D18-1041} creates in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$. $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively. While introducing new features in the architecture, \cite{D18-1041} also introduces new loss function which is sum of word-level weighted MT loss and losses of Domain-classifier in source side and target side; Adversarial Domain-classifier in source side. The authors have very good approach to seperate Domain-shared information flow and Domain-specify information flow and to determine their contributions to the inference thus mitigate the catastrophic interference occuring in training with backpropagation. \cite{D18-1039} has other idea to circumvent the catastrophic interference. They supposed that parameters of each domain specialized model could be generated from a universal parameter-generator. By introducing domain embedding (or Language embedding in the original paper), they could conduct domain related error in backpropagation toward domain embedding thus mitigates catastrophic interference between domains during training.

@@

\begin{table}[H]
\fyTodo{Too small ! Use abbrevs (En, Fr, ECB, EMEA, EPPS)}
  \resizebox{\columnwidth}{!}{
\begin{tabular}{ |l|l|l|l|l|l| }
\hline
Task & Corpora & Train & Dev & Test & Vocabulary \\ \hline
\multirow{5}{*}{$En \rightarrow Fr$} & European Medicines Agency & 1090568 & 1000 & 1000 & English: 30164, French: 30398\\
 & European Central Bank & 191960 & 1000 & 1000 &\\
 & European Parliament & 2005723 & 1000 & 1000 & \\ \hline
\multirow{5}{*}{$En \rightarrow Ge$} & European Medicines Agency & 1108752 & 1000 & 1000 & English: 30159, German: 30698\\
 & European Central Bank & 111174 & 1000 & 1000 &\\
 & European Parliament & 1920209 & 1000 & 1000 & \\ \hline
\hline
\end{tabular}
}
\caption{Corpora }
\label{table:Corpora}
\end{table}
\subsection{Baselines \label{ssec:baselines}}
\fyTodo{Describe baselines}

\subsection{NMT model configuration}
We evaluate our method on two state-of-the-art encoder-decoder architectures: a recurrent architecture based on attentional birectional Long Short-Term Memory (LSTM) cells and the Transformer architecture of \citet{NIPS2017_7181}.
More precisely, for the Transformer we use the same settings as described in \cite{NIPS2017_7181}. For the recurrent architecture, in order to fairly compare to the work of \citet{D18-1041}, we use a standard configuration with one layer bidirectional LSTM with a encoder hidden size of 1024; one attentional unidirectional LSTM layer with same hidden size in decoder; and word embeddings of size 512. The implementation can be found in OpenNMT-tf \cite{P17-4012}.
\fyTodo{Language pairs, BPE number, optimizer}
\fyTodo{Motivate the split}
For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions $[8,8,8,488]$, corresponding respectively to EMEA, EPPS, ECB domain and generic region. If sentence comes from domain $i$, regions for domains $j,\forall j \neq i$ will be masked as 0. We then use a dense layer of size 512 to fuse the active domain region and the generic region.

@@

Each iteration of algorithm \ref{algo:1}, the batch is doubled because it contains 2 types of word embedding for each sentence (using only generic region; using generic region and active domain region). The reason to do this is to minimize simultaneously both losses \\
$\theta^*_{s}=\displaystyle{\mathop{\argmin}_{\theta_s}}E_{(x,y) \sim D(\displaystyle{\mathop{\cup}_{i \in [1,..,d]}}C_{i})}[-log(p_{\theta_s}(y|x))]$ \\ 
$ \theta^*_{i}=\displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
$ 
We admit that this is not the optimal way. However, this procdedure make the training very simple.\\
We also take in account the problem of overfitting that could occur for small corpora. Indeed, real data are not always balanced for every domain. For instance, in our experiments, the size of corpora are very different where EPPS is largest with approximately 2 millions sentences, EMEA contains about 1 million sentences and ECB is the smallest with only 200 000 sentences for English French and 100 000 for English-German. If we train each corpora with same number of batches, ECB will be trained too oftenly, that easily leads to overfitting. Therefore, we perform backpropagation on corpora with frequencies according to sizes of corpora.
\subsection{Results}
\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 &  & 25.7 \\
 \hline
 EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 &  & 11.10 \\
 \hline
 EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 &  & 11.1 \\
 \hline
 ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 &  & 7.19 \\
 \hline
 Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(3) &  &  &  &  &  &  &  & \\
 \hline
 Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  & & \\
 \hline
\end{tabularx}}
\end{center}
\caption{BPE-detokenized BLEU score in language pair English - French. Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT. Sparse(2) uses generic tag for all of test sets. Sparse(3) uses tag predicted by pretrained domain classififer}
\label{tab:1}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & & & & & & & \\
 \hline
 EMEA-finetuned & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & \\
 \hline
 Sparse & & & & & & & \\
 \hline
 Domain Control & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & & & & & & & \\
 \hline 
\end{tabularx}}
\end{center}
\caption{BLEU score for Recurrent network on language pair English - German. Notations are the same as for figure~\ref{tab:3}}
\label{tab:4}
\end{figure}
\subsection{Word embedding analysis}

\section{Conclusions}
\fyTodo{natural continuations: go beyond words; learn the projection matrix at the word level; other ?}
